---
title: "Getting started with dkanr - retrieving information"
date: "`r Sys.Date()`"
output: 
 rmarkdown::html_vignette:
    toc: true
 rmarkdown::md_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to dkanr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

## Introduction to dkanr

The `dkanr` package is an R client to the [DKAN REST API](https://dkan.readthedocs.io/en/latest/apis/rest-api.html). Through the DKAN REST API, `dkanr` accesses the available catalog while providing CRUD functionalities.

This brief introduction shows you how to:

* Locate a dataset
* Access its metadata
* Download a .csv file of the data


## Setting Up Your Connection

### Connection without authentication

To set-up a connection without authentication, you just need the site URL. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(purrr)
library(dkanr)
library(dplyr)
dkanr_setup(url = 'https://datacatalog.worldbank.org/')
```

### Authenticated connection
If authentication is required, you will also need to specify your username and password.

```{r message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
dkanr_setup(url = 'http://demo.getdkan.com',
            username = 'my_username',
            password = 'my_password')
```

You can verify that you are successfully connected by printing your connection information.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
dkanr_settings()
```

## STEP 1: List all available datasets with `list_all_nodes()`

While exploring the offerings of the catalog, you can retrieve all the available datasets with a simple query.
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
resp <- list_all_nodes(filters = c(type = 'dataset'), as = 'df')
resp %>%
  select(nid, title, uri, type) %>%
  head(n = 10)
```

## STEP 2: Access metadata for a specific dataset node

Say we are interested in the following dataset: "Crowdsourced Price Data Collection Pilot". The metadata for this node can be retrieved as follows: 
1. Identify the dataset node ID
2. Use the node ID to retrieve the dataset metadata

### Identify the dataset node ID  

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
resp %>%
  filter(title == 'Crowdsourced Price Data Collection Pilot') %>%
  select(nid, title, uri, type)
```

### Use the node ID to retrieve the dataset metadata  

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
metadata <- retrieve_node(nid ='140177', as = 'list')
metadata
```

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# All metadata fields
names(metadata)[1:30]
```



```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Access specific metadata fields
metadata$title
```

## STEP 3: Access data for a specific resource node
Retrieving data file attached your dataset follows a similar logic:
1. Identify the resource node IDs
2. Use the resource node IDs to access metadata and URLs

### Retrieve the resource Node IDs

17 resources are attached to this dataset

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
get_resource_nids(metadata)
```

### Use the resource node ID to retrieve its metadata  

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
metadata_rs <- retrieve_node(nid ='140366', as = 'list')
metadata_rs
```


```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# All metadata fields
names(metadata_rs)[1:30]
```

### Retrieve the resource URL from the resource metadata

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
get_resource_url(metadata_rs)
```


## Querying data
Some data files may be directly queried through the API. Only data files that have been imported into the DKAN datastore can be queried through the API. 

### Check if a data file is available on the DKAN datastore

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
ds_is_available(metadata_rs)
```

### Retrieve data from the datastore via the API

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
ds_search_all(resource_id = metadata_rs$uuid, as = 'df') %>%
  select(item_name, item_code, price)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
temp <- ds_search(resource_id = metadata_rs$uuid)
dplyr::bind_rows(temp) %>%
  select(item_name, item_code, price)
```
